The Effect of Media Frames on Issue Attitudes

### 2021-10-04 paper/main.tex
Updates from Overleaf

While news slant is a very general concept, and sometimes operationalised as similarity to specific parties' communication (\cite{Gentzkow2010}) or issue positions/tonality (\cite{Spirig2020}), framing theory provides a specific expectation about which treatment should affect which issue attitudes how. As \citeauthor{Nelson1997} write: "framing is the process by which a communication source, such as a news organisation, defines and constructs a political issue or public controversy" (\citeyear[567]{Nelson1997}). A "frame" is then the operating unit of this process, defining the issue at hand in a specific way. Different forms of frames have been conceptualised. When I talk about 'frames' here, I refer to \textit{emphasis frames}. These frames emphasise certain topics in relation to an issue, guiding the recipient to think about the issue with those considerations in mind that are promoted by the frame (\cite[153f]{Leeper2020}). For example, issues of increasing welfare contributions might be discussed with reference to inequality and providing chances to the poor, or by mentioning that higher welfare benefits might result in higher taxes. Individuals' support of a statement will differ dependent on the frame presented (\cite{sniderman2004structure}).

@@

Figure \ref{fig:cogStor} visualises this logic. A given issue is associated with an attitude or evaluation $y$. The issue is related to a number of different considerations, each of which carries an associated evaluation $x_i$, e.g. that the prevention of overcrowded hospitals is a good thing. Each of these considerations is more or less associated with the issue. Based on the strength of this association $w_i$, a given consideration's evaluation is more or less reflected in the issue evaluation $y$, which is equal to the weighted sum of considerations. Emphasis framing affects the issue attitude $y$ by changing the weights of different considerations\footnote{Note that this concept of framing is identical to second-level or attribute agenda setting (\cite{Lopezescobar2017, McCombs2000}; see \cite[174]{Mclaren2018} for a discussion of the overlap of these literatures). I will stick to the framing perspective here as the literature is focused on issues rather than candidates' attributes, and the conceptualisation as frame allows to form expectations about citizens' reactions to changing news frames.}.

This operationalisation allows us to generate specific expectations about how frames in the news media affect respondents' issue opinions. When presented with an emphasis frame on a given issue, the respondents' cognitive association of the issue with the emphasised consideration should increase, thus changing the evaluation of the issue by increasing the importance of the consideration. If the consideration is positive (towards the issue), the attitude towards the issue should improve, and the opposite when it is negative. \medskip

\textbf{H2 (emphasis framing):} If a newspaper emphasises more positive/negative frames of an issue, the readers' attitude about the issue should become more positive/negative.

\begin{center}
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{paper/vis/CognitiveStorage.png}
    \caption{The cognitive evaluation of an issue $y$ is determined by the strength of association $w_i$ with other concepts with existing evaluations $x_i$.}
    \label{fig:cogStor}
\end{figure}
\end{center}

@@

[describe context of German election 2017 and dominance of migration topic]

@@

To measure the dependent variable (issue attitudes) and differentiate the treatment groups (newspaper readers), I employ panel data from the German Longitudinal Election Study (GLES, \cite{GLES2019LongTermTracking}). The panel consists of 15 waves, of which 6 (4) contain variables on newspaper consumption and immigration (integration) attitudes. In each wave 9,000-15,000 respondents were interviewed, which allows the precise estimation of effects of media consumption.

Figure \ref{fig:issues} shows the two main dependent variables for each group of readers. Note that higher values on the immigration variable indicate more restrictive attitudes, while higher values on the integration variable are most liberal. As can be seen, readers of the tabloid 'Bild' differ strongly from consumers of other outlets, with far more restrictive attitudes. Both variables seem inversely correlated (as expected), with values moving towards the more liberal pole until autumn, where the trend is reversed. The development of issue importance across time can be found in Appendix \ref{app:importance}.

@@

To measure media attention to immigration, I collected 2.5 million newspaper articles from the most important German broadsheets \textit{Frankfurter Allgemeine Zeitung (FAZ), Spiegel Online (SPON), Süddeutsche Zeitung (SZ), Die Tageszeitung (TAZ), Die Welt} and the major German tabloid \textit{Bild} for the period 2016-2019. To identify whether an article is about migration, I train a BERT deep-learning classifier.

As migration content is rare, and my training data needs to be somewhat balanced, I first construct a dictionary of terms related to migration. I use dictionary extension based on German Glove\footnote{Downloadable from \url{https://deepset.ai/german-word-embeddings}.} to construct a comprehensive dictionary, and apply it to the articles. Based on the relative share of migration words in an article, I draw a stratified sample of 300 articles (100 from the third with fewest migration terms, 100 from the third with most migration terms and 100 from the third in between) for each paper for a total of 1,800 papers. Then, a student assistant\footnote{Thank you Robin.} hand-coded these articles, assessing whether their main topic is related to migration.

Using this sample, a BERT deep-learning classifier\footnote{\url{https://huggingface.co/bert-base-german-cased}} is fine-tuned on a subset of 1,400 annotated articles. After fine-tuning, the model correctly classifies 95.5\% of the test set (F1: 0.94, recall: 0.93, precision: 0.95). This classifier is then used to annotate all 2.5 million newspaper articles. For the time frame of relevance for this study (the entire year of 2017), out of the total of 388,617 articles, 13,589 (3.5\%) are identified to treat the issue of migration.

@@

Figure \ref{fig:salience} shows the share of migration articles in each newspaper. While the first few months of 2017 show rather high salience of the issue, attention decreases until May, before the share of articles covering this issue increases again in all newspapers except Bild, which shows a stark decrease in July and maintains this level afterward. Most newspapers show similar levels of attention to the issue between 2.5\% and 6\%, with the exception of SZ, which moves between 5\% and 12.5\%.

@@

To measure news frames, I estimate a structural topic model with 60 topics for the nearly 90,000 migration articles in the full period (2016-2019)\footnote{The }. Topic prevalence is estimated as a function of the release date of the article, as well as the paper it has been published in. The 60 topics' most predictive words and ten most representative articles were assessed to assess the content of each topic, and the most frequent and relevant topics that allow to form clear hypotheses regarding their effect on immigration attitudes are selected.

With this procedure, the following seven 'large' migration frames were selected:

\begin{enumerate}
    \item Capital crime (sexual assault/rape/murder) committed by refugees, 
    \item general petty crime by refugees and issues in refugee living quarters in Germany,
    \item refugee numbers,
    \item labour market needs for and integration of refugees,
    \item deportations,
    \item internment camps (e.g. Moria),
    \item drownings in the Mediterranean.
\end{enumerate}

I expect both crime frames and the refugee numbers to have negative effects on attitudes towards migrants (as connected to negative considerations regarding immigration), while deportations, internment camps, and drownings in the Mediterranean sea should have positive effects, framing immigration either as a labour market need or a humanitarian necessity. The labour market frame should show mixed results, given that labour market integration might be perceived as threatening or addressing needs in the German economy. Figure \ref{fig:frames} shows the prevalence of all frames across time for each newspaper. It can be seen that substantial variation exists in the relative emphasis of migration frames both within and across newspapers.

@@

\subsection{Correlation of DiD with frame prevalence}
% show correlation of DiD with prevalence of media frames (STM)
\begin{figure}[!ht]
    \centering
    \includegraphics[width=\textwidth]{paper/vis/effectplot_frames.png}
    \caption{Correlation of opinion shifts (DiDs) with media attention or shifts in attention (DiD's) for different specifications.}
    \label{fig:did_corr}
\end{figure}

@@

\subsection{Descriptives on issue importance across time and newspaper consumption}\label{app:importance}

[add description and plot issue importance]

>>>>

While news slant is a very general concept, and sometimes operationalised as similarity to specific parties' communication (\cite{Gentzkow2010}) or issue positions/tonality (\cite{Spirig2020}), framing theory provides a specific expectation about which treatment should affect which issue attitudes how. As \citeauthor{Nelson1997} write: "framing is the process by which a communication source, such as a news organisation, defines and constructs a political issue or public controversy" (\citeyear[567]{Nelson1997}). A "frame" is then the operating unit of this process, defining the issue at hand in a specific way. Different forms of frames have been conceptualised. When I talk about 'frames' here, I refer to \textit{emphasis frames}. These frames emphasise certain topics in relation to an issue, guiding the recipient to think about the issue with those considerations in mind that are promoted by the frame (\cite[153f]{Leeper2020}). For example, issues of increasing welfare benefits might be discussed with reference to inequality and providing chances to the poor, or by mentioning that higher welfare benefits might result in higher taxes. Individuals' support of a statement will differ dependent on the frame presented (\cite{sniderman2004structure}).

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{paper/vis/CognitiveStorage.png}
    \caption{The cognitive evaluation of an issue $y$ is determined by the strength of association $w_i$ with other concepts with existing evaluations $x_i$.}
    \label{fig:cogStor}
\end{figure}

@@

Figure \ref{fig:cogStor} visualises this logic. A given issue is associated with an attitude $y$. The issue is related to a number of different considerations, each of which carries an associated evaluation $x_i$, e.g. that the prevention of overcrowded hospitals is a good thing. Each of these considerations is more or less associated with the issue. Based on the strength of this association $w_i$, a given consideration's evaluation is more or less reflected in the issue evaluation $y$, which is equal to the weighted sum of considerations. Emphasis framing affects the issue attitude $y$ by changing the weights of different considerations\footnote{Note that this concept of framing is identical to second-level or attribute agenda setting (\cite{Lopezescobar2017, McCombs2000}; see \cite[174]{Mclaren2018} for a discussion of the overlap of these literatures). I will stick to the framing perspective here as the literature is focused on issues rather than candidates' attributes, and the conceptualisation as frame allows to form expectations about citizens' reactions to changing news frames.}.

This operationalisation allows me to generate specific expectations about how frames in the news media affect respondents' issue opinions. When presented with an emphasis frame on a given issue, the respondents' cognitive association of the issue with the emphasised consideration should increase, thus changing the evaluation of the issue by increasing the importance of the consideration. If the consideration is positive (towards the issue), the attitude towards the issue should improve, and the opposite when it is negative. \medskip

\textbf{Emphasis framing hypothesis:} If a newspaper emphasises more positive/negative frames of an issue, the readers' attitude about the issue becomes more positive/negative.

@@

To test this hypothesis, I analyse media and opinion data on the migration issue from the German national election in 2017. Following a period of increased refugee arrivals in Europe in the summer of 2015, the election was dominated by the migration topic and the related upcoming of the radical-right populist challenger party \textit{AfD}. The major two parties dominating German post-war politics, \textit{SPD} and \textit{CDU/CSU}, had to face severe losses following their 'grand coalition' (\cite{Bieber2021, Wessels2021}).

This strong (although still variant, see figure \ref{fig:salience}) salience of the migration issue both in the media as well as in citizens subjective evaluations presents an ideal case to study the effects of shifting media framing on issue attitudes, as continuous coverage on and public debate of the issue are a given.

@@

To measure the dependent variable (issue attitudes) and differentiate the treatment groups (newspaper readers), I employ panel data from the German Longitudinal Election Study (\cite{GLES2019LongTermTracking}). The panel consists of 15 waves, of which 6 (4) contain variables on newspaper consumption and immigration (integration) attitudes. In each wave 9,000-15,000 respondents were interviewed, which allows the precise estimation of effects of media consumption.

Figure \ref{fig:issues} shows the two main dependent variables for each group of readers. Note that higher values indicate more restrictive attitudes towards immigration/integration (see plot subtitles for more information). As can be seen, readers of the tabloid 'Bild' differ strongly from consumers of other outlets, with far more restrictive attitudes. Both variables seem correlated (as expected), with values moving towards the more liberal pole until autumn, where the trend is reversed.

@@

@@

To measure media attention to immigration, I collected 2.5 million newspaper articles from the most important German broadsheets \textit{Frankfurter Allgemeine Zeitung (FAZ), Spiegel Online (SPON), Süddeutsche Zeitung (SZ), Die Tageszeitung (TAZ), Die Welt} and the major German tabloid \textit{Bild} for the period 2013-2019. To identify whether an article is about migration, I train a BERT deep-learning classifier.

As migration content is rare, and my training data needs to be somewhat balanced, I first construct a dictionary of terms related to migration. I use dictionary extension based on German GloVe word-embeddings\footnote{Downloadable from \url{https://deepset.ai/german-word-embeddings}.} to construct a comprehensive dictionary, and apply it to the articles. Based on the relative share of migration words in an article, I draw a stratified sample of 300 articles (100 from the third with fewest migration terms, 100 from the third with most migration terms and 100 from the third in between) for each newspaper for a total of 1,800 papers. Then, a student assistant\footnote{Thank you Robin.} hand-coded these articles, assessing whether their main topic is related to migration.

Using this sample, a BERT deep-learning classifier\footnote{\url{https://huggingface.co/bert-base-german-cased}} is fine-tuned on a subset of 1,400 annotated articles. After fine-tuning, the model correctly classifies 95.5\% of the test set (F1: 0.94, recall: 0.93, precision: 0.95). This classifier is then used to annotate all 2.5 million newspaper articles. For the time frame of relevance for this study (the entire year of 2017), out of the total of 388,617 articles, 13,589 (3.5\%) are identified to treat the issue of migration\footnote{Note that the classifier, as well as the following topic model, was trained on a corpus of documents spanning 2013-2019. This was largely a decision to maximise possible applications for the resulting dataset, but also enables to compare issue attention and frames across time for validity checks.}.

Figure \ref{fig:salience} shows the share of migration articles in each newspaper. While the first few months of 2017 show rather high salience of the issue, attention decreases until May, before the share of articles covering this issue increases again in all newspapers except \textit{Bild}, which shows a stark decrease in July and maintains this level afterward. Most newspapers show similar levels of attention to the issue between 2.5\% and 6\%, with the exception of \textit{SZ}, which moves between 5\% and 12.5\%.

@@

To measure news frames, I estimate a structural topic model with 60 topics for the nearly 90,000 migration articles in the full period (2016-2019)\footnote{The topics' prevalence along with the most important words can be assessed in appendix \ref{app:stm}.}. The number of topics was chosen to strike a balance between the computational and human resources necessary to estimate and annotate the topics and finding appropriate topics for the large corpus of nearly 90,000 migration articles. Topic prevalence is estimated as a function of the release date of the article, as well as the newspaper it has been published in. The 60 topics' most predictive words and ten most representative articles were assessed to identify and label the content of each topic, and the most frequent and relevant topics that allow to form clear hypotheses regarding their effect on immigration attitudes are selected.

With this procedure, the following seven 'large' migration frames were selected (number of topic added corresponds to the graph in appendix \ref{app:stm}):

\begin{enumerate}
    \item Capital crime (sexual assault/rape/murder) committed by refugees (topic 48), 
    \item illegal entry and petty crime (topic 25),
    \item refugee numbers (topic 32),
    \item labour market needs for and integration of refugees (topic 11),
    \item deportations (topic 44),
    \item internment camps (e.g. Moria; topic 27),
    \item drownings in the Mediterranean (topic 49).
\end{enumerate}

I expect the capital crime, illegal entry/petty crime, and the refugee numbers frames to have negative effects on attitudes towards migrants (as connected to negative considerations regarding immigration), while deportations, internment camps, and drownings in the Mediterranean sea should have positive effects, framing immigration as a humanitarian issue. The labour market frame should show mixed results, given that labour market integration might be perceived as threatening or addressing needs in the German economy. Figure \ref{fig:frames} shows the prevalence of all frames across time for each newspaper. It can be seen that substantial variation exists in the relative emphasis of migration frames both within and across newspapers.

@@

\subsection{Correlation of DiD with frame prevalence}
% show correlation of DiD with prevalence of media frames (STM)
\begin{figure}[!ht]
    \centering
    \includegraphics[width=\textwidth]{paper/vis/effectplot_frames_did.png}
    \caption{Correlation of opinion shifts (DiDs) with media attention or shifts in attention (DiD's) for different specifications.}
    \label{fig:did_corr}
\end{figure}

@@

\subsection{Extended migration dictionary}\label{app:dictionary}

The extended migration dictionary for the pre-sampling of articles on migration consists of 35 terms (matching both lower and upper case): \textit{asyl, asylant, asylanten, asylbewerber, asylbewerberin, auslander, einwanderer, einwandernde, einwanderung, migration, immigration, schutzsuchende, visa, visum, zuwanderung, zuwanderer, zuwandernde, ausländer, ausländerinnen, flüchtling, flüchtende, flüchtlinge, flüchtlingen, migranten, vertriebene, asylantinnen, asylbewerberinnen, auslanderinnen, einwandererin, einwandererinnen, fluchtling, fluchtende, fluchtlinge, schutzbedurftige, schutzbedürftige.}
